---
title: "Lateral analysis"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading data

```{r}
library(pacman)
p_load(tidyverse, ggsignif, multcomp, broom, emmeans, DescTools)
setwd("~/Documents/Systemic ESYT1 project locomotion, behavior, survival, anatomical data/Locomotion analysis cleaned/") 
df<- read.table("statistics.csv", sep=",", header=T)


```
Assign time point

```{r}
df2 <- df
df2$time_point <- 0
df2$time_point <- ifelse(grepl('p49_', df2$name), 49, df2$time_point)
df2$time_point <- ifelse(grepl('p56_', df2$name), 56, df2$time_point)
df2$time_point <- ifelse(grepl('p63_', df2$name), 63, df2$time_point)
df2$time_point <- ifelse(grepl('p70_', df2$name), 70, df2$time_point)
df2$time_point <- ifelse(grepl('p77_', df2$name), 77, df2$time_point)
df2$time_point <- ifelse(grepl('p84_', df2$name), 84, df2$time_point)
df2$time_point <- ifelse(grepl('p91_', df2$name), 91, df2$time_point)
df2$time_point <- ifelse(grepl('p98_', df2$name), 98, df2$time_point)
df2$time_point <- ifelse(grepl('p105_', df2$name), 105, df2$time_point)
df2$time_point <- ifelse(grepl('p112_', df2$name), 112, df2$time_point)

```



Give each mouse own number
```{r}

df2$mouse_number <- 0

#SOD1En1cre
df2$mouse_number <- ifelse(grepl('736_', df$name), '736', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('589_', df$name), '589', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('603_', df$name), '603', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('607_', df$name), '607', df2$mouse_number)
#En1re
df2$mouse_number <- ifelse(grepl('561_', df$name), '561', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('694_', df$name), '694', df2$mouse_number) 
df2$mouse_number <- ifelse(grepl('696_', df$name), '696', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('698_', df$name), '698', df2$mouse_number)
#SOD1
df2$mouse_number <- ifelse(grepl('730_', df$name), '730', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('734_', df$name), '734', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('738_', df$name), '738', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('702_', df$name), '702', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('627_', df$name), '627', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('625_', df$name), '625', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('615_', df$name), '615', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('788_', df$name), '788', df2$mouse_number)
#WT
df2$mouse_number <- ifelse(grepl('692_', df$name), '692', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('700_', df$name), '700', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('706_', df$name), '706', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('708_', df$name), '708', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('553_', df$name), '553', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('555_', df$name), '555', df2$mouse_number)
```

Assign genotypes
```{r}
df2$Genotype <- 0


df2$Genotype <- ifelse(df2$mouse_number==702, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==730, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==734, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==738, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==625, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==627, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==615, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==788, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==700, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==692, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==706, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==708, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==553, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==555, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==694, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==698, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==696, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==561, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==736, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==589, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==603, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==607, 'SOD1En1cre', df2$Genotype)

class(df2$Genotype)
```


Reduce dataset to the 6 SOD1 that were injected with GFP

```{r}
df2<-df2[!(df2$mouse_number=='625' | df2$mouse_number=='627'),] 
```

Compute mean for each mouse for each timepoint

```{r}
df_means_mouse_by_Genotype <- df2 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(hip_ang = mean(hip_ang), knee_ang = mean(knee_ang), ankle_ang = mean(ankle_ang), foot_ang = mean(foot_ang)) %>% 
  arrange(Genotype)
```

Create dataset with data from onset of locomotor phenotype

Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the WT mice
```{r}
#color_palette inspired colorpalette
color_palette = c("#5B1A18","#FD6467",  "#F1BB7B")

```

Removing En1
```{r}
df2<-df2[!(df2$Genotype=='En1cre'),] 
```

Preparing the ggsignif function to use Dunnet's p-values for bar plots
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}


```

Analysis of p112

```{r}
p112 <- subset(df_means_mouse_by_Genotype, time_point==112) #Mean values
p112_all <- subset(df2, time_point==112) #Video triplicates
```

one-way anova at p112
 
```{r}

#Ordering Genotype variables by manipulation so SOD1 becomes 'control' in Dunnet post-hoc tests
p112_all$Genotype <- factor(p112_all$Genotype, levels = c("SOD1","SOD1En1cre", "WT", "En1cre"))

#foot angle
anova_foot_1way <- aov(foot_ang ~ Genotype, data = p112_all)
summary(anova_foot_1way)

DunnettTest(x=p112_all$foot_ang, g=p112_all$Genotype)

#ankle
anova_ankle_1way <- aov(ankle_ang ~ Genotype, data = p112_all)
summary(anova_ankle_1way)

DunnettTest(x=p112_all$ankle_ang, g=p112_all$Genotype)

#knee
anova_knee_1way <- aov(knee_ang ~ Genotype, data = p112_all)
summary(anova_knee_1way)

DunnettTest(x=p112_all$knee_ang, g=p112_all$Genotype)

#hip
anova_hip_1way <- aov(hip_ang ~ Genotype, data = p112_all)
summary(anova_hip_1way)

DunnettTest(x=p112_all$hip_ang, g=p112_all$Genotype)


summary(anova_hip_1way)
```



Bar plots p112_all

```{r}


#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("WT", "En1cre", "SOD1", "SOD1En1cre"))
p112_all$Genotype <- factor(p112_all$Genotype, levels = c("WT", "En1cre", "SOD1", "SOD1En1cre"))


#foot
Dunnet<- glht(anova_foot_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= p112_all, aes(x=Genotype, y=foot_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+

  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Foot angle")+
    geom_jitter(data=p112, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(), text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(70,62,66)
  )

#ankle
Dunnet<- glht(anova_ankle_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= p112_all, aes(x=Genotype, y=ankle_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 

  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Ankle angle")+
    geom_jitter(data=p112, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(47,41,44)
  )

#knee
Dunnet<- glht(anova_knee_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= p112_all, aes(x=Genotype, y=knee_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 

  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Knee angle")+
    geom_jitter(data=p112, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(23,20,21.5)
  )

#hip
Dunnet<- glht(anova_hip_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= p112_all, aes(x=Genotype, y=hip_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+

  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Hip angle")+
    geom_jitter(data=p112, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(60,50,55)
  )


```
Analysis of onset



Analysis of onset


create subset of videos from onset of locomotor phenotype (when mouse could not keep up with treadmill belt speed of 20 cm/s). If mouse did not show onset at any point, onset is set to p112, except for WT and En1cre, they are set to p63. for mice where they have e.g had 2 trials at 20 cm and then one trial at 15 cm, the 'onset' has been counted as the next session where all trials were under 20 cm in speed.

triplicates
```{r}
class(df2$mouse_number)

onset_730 <-df2 %>% filter(mouse_number =='730' & time_point == 84)
onset_734 <- df2%>% subset((mouse_number =='734' & time_point == 84))
onset_736 <- df2%>% subset((mouse_number =='736' & time_point == 112))
onset_738 <- df2%>% subset((mouse_number =='738' & time_point == 91))
onset_788 <- df2%>% subset((mouse_number =='788' & time_point == 84))
onset_702 <- df2%>% subset((mouse_number =='702' & time_point == 98))
onset_615 <- df2%>% subset((mouse_number =='615' & time_point == 91))
onset_589 <- df2%>% subset((mouse_number =='589' & time_point == 112))
onset_607 <- df2%>% subset((mouse_number =='607' & time_point == 112))
onset_603 <- df2%>% subset((mouse_number =='603' & time_point == 112))

onset_WT <- df2%>% subset((Genotype == 'WT' & time_point == 63))
#onset_En1cre <- df2%>% subset((Genotype == 'En1cre' & time_point == 63))


#did not include En1cre as we dod not collect data from them until p112, so do not think they should be included for analysis
onset_all <- rbind(onset_730, onset_734, onset_736, onset_738, onset_788, onset_702, onset_615, onset_589, onset_607, onset_603, onset_WT)
```

Means


for mean
```{r}

onset_730 <-df_means_mouse_by_Genotype%>% filter(mouse_number =='730' & time_point == 84)
onset_734 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='734' & time_point == 84))
onset_736 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='736' & time_point == 112))
onset_738 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='738' & time_point == 91))
onset_788 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='788' & time_point == 84))
onset_702 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='702' & time_point == 98))
onset_615 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='615' & time_point == 91))
onset_589 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='589' & time_point == 112))
onset_607 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='607' & time_point == 112))
onset_603 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='603' & time_point == 112))

onset_WT <- df_means_mouse_by_Genotype%>% subset((Genotype == 'WT' & time_point == 63))
#onset_En1cre <- df_means_mouse_by_Genotype%>% subset((Genotype == 'En1cre' & time_point == 63))


#did not include En1cre as we dod not collect data from them until p112, so do not think they should be included for analysis
onset_mean <- rbind(onset_730, onset_734, onset_736, onset_738, onset_788, onset_702, onset_615, onset_589, onset_607, onset_603, onset_WT)

```

one-way anova at onset

```{r}

#Ordering Genotype variables by manipulation so SOD1 becomes 'control' in Dunnet post-hoc tests
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "WT", "En1cre", "SOD1En1cre"))

#foot angle
anova_foot_1way <- aov(foot_ang ~ Genotype, data = onset_all)
summary(anova_foot_1way)

DunnettTest(x=onset_all$foot_ang, g=onset_all$Genotype)

#ankle
anova_ankle_1way <- aov(ankle_ang ~ Genotype, data = onset_all)
summary(anova_ankle_1way)
DunnettTest(x=onset_all$ankle_ang, g=onset_all$Genotype)

#knee
anova_knee_1way <- aov(knee_ang ~ Genotype, data = onset_all)
summary(anova_knee_1way)
DunnettTest(x=onset_all$knee_ang, g=onset_all$Genotype)

#hip
anova_hip_1way <- aov(hip_ang ~ Genotype, data = onset_all)
summary(anova_hip_1way)
DunnettTest(x=onset_all$hip_ang, g=onset_all$Genotype) #Dunnet's test

```


Bar plots onset only

```{r}

#Re-ordering so the experimental groups appear in order of control level
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("WT", "En1cre","SOD1", "SOD1En1cre"))

#foot
Dunnet<- glht(anova_foot_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= onset_all, aes(x=Genotype, y=foot_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Foot angle")+
  geom_jitter(data= onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=onset_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(59,63,67)
  )

#ankle
Dunnet<- glht(anova_ankle_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= onset_all, aes(x=Genotype, y=ankle_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Ankle angle")+
  geom_jitter(data= onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=onset_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(41, 44, 47)
  )

#knee

Dunnet<- glht(anova_knee_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= onset_all, aes(x=Genotype, y=knee_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Knee angle")+
  geom_jitter(data= onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=onset_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(), text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(20,21.5,23)
  )

#hip
Dunnet<- glht(anova_hip_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= onset_all, aes(x=Genotype, y=hip_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Hip angle")+
  geom_jitter(data= onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=onset_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(48,51,54)
  )

```
Longitudinal plots


Hip Angle
```{r}

#Re-ordering so the experimental groups appear in order of control level
df2$Genotype <- factor(df2$Genotype, levels = c("WT", "En1cre", "SOD1", "SOD1En1cre"))

ggplot(df2, aes(x=time_point, y= hip_ang, color=Genotype)) +
  geom_point(stat = "summary",fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary",fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Hip Angle")

```

knee angle
```{r}
ggplot(df2, aes(x=time_point, y= knee_ang, color=Genotype)) +
  geom_point(stat = "summary",fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary",fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Knee Angle")

```

Ankle angle
```{r}
ggplot(df2, aes(x=time_point, y= ankle_ang, color=Genotype)) +
  geom_point(stat = "summary",fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary",fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Ankle Angle")

```

foot angle
```{r}
ggplot(df2, aes(x=time_point, y= foot_ang, color=Genotype)) +
  geom_point(stat = "summary",fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary",fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Foot Angle")

```

Analysis across timepoints

```{r}
df2$time_point <- as.factor(df2$time_point) #independent variables have to be recoded as factors to allow 'stratification' by timepoint for two-way anova

#refactoring genotype to set SOD1 as 'control'
df2$Genotype <- factor(df2$Genotype, levels = c("SOD1", "WT", "En1cre", "SOD1En1cre"))
```

two-way Anovas and post-hoc
```{r}

#foot
anova_foot_2way <- aov(foot_ang ~ Genotype*time_point, data = df2)
summary(anova_foot_2way)
emmeans(anova_foot_2way,specs = trt.vs.ctrl~ Genotype|time_point)

### Ankle
anova_ankle_2way <- aov(ankle_ang ~ Genotype*time_point, data = df2)
summary(anova_ankle_2way)

emmeans(anova_ankle_2way,specs = trt.vs.ctrl~ Genotype|time_point)

### Knee
anova_knee_2way <- aov(knee_ang ~ Genotype*time_point, data = df2)
summary(anova_knee_2way)

emmeans(anova_knee_2way,specs = trt.vs.ctrl~ Genotype|time_point)


###Hip
anova_hip_2way <- aov(hip_ang ~ Genotype*time_point, data = df2)
summary(anova_hip_2way)

emmeans(anova_hip_2way,specs = trt.vs.ctrl~ Genotype|time_point)


```