---
title: "timepoint analysis"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(pacman)
p_load(tidyverse, ggsignif,multcomp, broom, emmeans, DescTools)
setwd("~/Documents/Systemic ESYT1 project locomotion, behavior, survival, anatomical data/Locomotion analysis cleaned/") 
df<- read.table("speedProfile_final.csv", sep="\t", header=T)


```
Assign time point

```{r}
df2 <- df
df2$time_point <- 0
df2$time_point <- ifelse(grepl('p49_', df2$Name), 49, df2$time_point)
df2$time_point <- ifelse(grepl('p56_', df2$Name), 56, df2$time_point)
df2$time_point <- ifelse(grepl('p63_', df2$Name), 63, df2$time_point)
df2$time_point <- ifelse(grepl('p70_', df2$Name), 70, df2$time_point)
df2$time_point <- ifelse(grepl('p77_', df2$Name), 77, df2$time_point)
df2$time_point <- ifelse(grepl('p84_', df2$Name), 84, df2$time_point)
df2$time_point <- ifelse(grepl('p91_', df2$Name), 91, df2$time_point)
df2$time_point <- ifelse(grepl('p98_', df2$Name), 98, df2$time_point)
df2$time_point <- ifelse(grepl('p105_', df2$Name), 105, df2$time_point)
df2$time_point <- ifelse(grepl('p112_', df2$Name), 112, df2$time_point)

```



Give each mouse own number
```{r}

df2$mouse_number <- 0

#SOD1En1cre
df2$mouse_number <- ifelse(grepl('736_', df$Name), '736', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('589_', df$Name), '589', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('603_', df$Name), '603', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('607_', df$Name), '607', df2$mouse_number)
#En1re
df2$mouse_number <- ifelse(grepl('561_', df$Name), '561', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('694_', df$Name), '694', df2$mouse_number) 
df2$mouse_number <- ifelse(grepl('696_', df$Name), '696', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('698_', df$Name), '698', df2$mouse_number)
#SOD1
df2$mouse_number <- ifelse(grepl('730_', df$Name), '730', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('734_', df$Name), '734', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('738_', df$Name), '738', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('702_', df$Name), '702', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('625_', df$Name), '625', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('627_', df$Name), '627', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('615_', df$Name), '615', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('788_', df$Name), '788', df2$mouse_number)
#WT
df2$mouse_number <- ifelse(grepl('692_', df$Name), '692', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('700_', df$Name), '700', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('706_', df$Name), '706', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('708_', df$Name), '708', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('553_', df$Name), '553', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('555_', df$Name), '555', df2$mouse_number)
```

Assign genotypes
```{r}
df2$Genotype <- 0


df2$Genotype <- ifelse(df2$mouse_number==702, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==730, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==734, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==738, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==625, 'SOD1', df2$Genotype) #SOD/WT injected with hEsyt1
df2$Genotype <- ifelse(df2$mouse_number==627, 'SOD1', df2$Genotype) #SOD/WT injected with hEsyt1
df2$Genotype <- ifelse(df2$mouse_number==615, 'SOD1', df2$Genotype) #SOD/En injected with GFP
df2$Genotype <- ifelse(df2$mouse_number==788, 'SOD1', df2$Genotype) #SOD/En injected with GFP
df2$Genotype <- ifelse(df2$mouse_number==700, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==692, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==706, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==708, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==553, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==555, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==694, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==698, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==696, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==561, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==736, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==589, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==603, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==607, 'SOD1En1cre', df2$Genotype)

```


Reduce dataset to the 6 SOD1 that were injected with GFP

```{r}
df2<-df2[!(df2$mouse_number=='625' | df2$mouse_number=='627'),] 
```

Generating mean cadence and stride for each observation
```{r}
#Calculating mean cadence
df2$MeanCadence <- rowMeans(df2[,c("fLCad.","fRCad", "hLCad.", "hRCad.")])

#Calculating mean stride
df2$MeanStride <- rowMeans(df2[,c("hLStride","hRStride", "fLStride", "fRStride")])
```


Creating datasets based on timepoint

```{r}
p49 <- newdata <- subset(df2, time_point==49)
p56 <- newdata <- subset(df2, time_point==56)
p63 <- newdata <- subset(df2, time_point==63)
p70 <- newdata <- subset(df2, time_point==70)
p77 <- newdata <- subset(df2, time_point==77)
p84 <- newdata <- subset(df2, time_point==84)
p91 <- newdata <- subset(df2, time_point==91)
p98 <- newdata <- subset(df2, time_point==98)
p105 <- newdata <- subset(df2, time_point==105)
p112 <- newdata <- subset(df2, time_point==112)
```

#Section used for creating excel files over genotypes and indivdiual mice, including mean stats for speed, acceleration, cadence and stride
Subset p49 by Genotype
```{r}
SOD1_p49 <- newdata <- subset(p49, Genotype=='SOD1')
SOD1En1cre_p49 <- newdata <- subset(p49, Genotype=='SOD1En1cre')
En1cre_p49 <- newdata <- subset(p49, Genotype=='En1cre')
WT_p49 <- newdata <- subset(p49, Genotype=='WT')
```

Subset p56 by Genotype
```{r}
SOD1_p56 <- newdata <- subset(p56, Genotype=='SOD1')
SOD1En1cre_p56 <- newdata <- subset(p56, Genotype=='SOD1En1cre')
En1cre_p56 <- newdata <- subset(p56, Genotype=='En1cre')
WT_p56 <- newdata <- subset(p56, Genotype=='WT')
```

Subset p63 by Genotype
```{r}
SOD1_p63 <- newdata <- subset(p63, Genotype=='SOD1')
SOD1En1cre_p63 <- newdata <- subset(p63, Genotype=='SOD1En1cre')
En1cre_p63 <- newdata <- subset(p63, Genotype=='En1cre')
WT_p63 <- newdata <- subset(p63, Genotype=='WT')
```

Subset p70 by Genotype
```{r}
SOD1_p70 <- newdata <- subset(p70, Genotype=='SOD1')
SOD1En1cre_p70 <- newdata <- subset(p70, Genotype=='SOD1En1cre')
En1cre_p70 <- newdata <- subset(p70, Genotype=='En1cre')
WT_p70 <- newdata <- subset(p70, Genotype=='WT')
```

Subset p77 by Genotype
```{r}
SOD1_p77 <- newdata <- subset(p77, Genotype=='SOD1')
SOD1En1cre_p77 <- newdata <- subset(p77, Genotype=='SOD1En1cre')
En1cre_p77 <- newdata <- subset(p77, Genotype=='En1cre')
WT_p77 <- newdata <- subset(p77, Genotype=='WT')
```

Subset p84 by Genotype
```{r}
SOD1_p84 <- newdata <- subset(p84, Genotype=='SOD1')
SOD1En1cre_p84 <- newdata <- subset(p84, Genotype=='SOD1En1cre')
En1cre_p84 <- newdata <- subset(p84, Genotype=='En1cre')
WT_p84 <- newdata <- subset(p84, Genotype=='WT')
```

Subset p91 by Genotype
```{r}
SOD1_p91 <- newdata <- subset(p91, Genotype=='SOD1')
SOD1En1cre_p91 <- newdata <- subset(p91, Genotype=='SOD1En1cre')
En1cre_p91 <- newdata <- subset(p91, Genotype=='En1cre')
WT_p91 <- newdata <- subset(p91, Genotype=='WT')
```

Subset p98 by Genotype
```{r}
SOD1_p98 <- newdata <- subset(p98, Genotype=='SOD1')
SOD1En1cre_p98 <- newdata <- subset(p98, Genotype=='SOD1En1cre')
En1cre_p98 <- newdata <- subset(p98, Genotype=='En1cre')
WT_p98 <- newdata <- subset(p98, Genotype=='WT')
```

Subset p105 by Genotype
```{r}
SOD1_p105 <- newdata <- subset(p105, Genotype=='SOD1')
SOD1En1cre_p105 <- newdata <- subset(p105, Genotype=='SOD1En1cre')
En1cre_p105 <- newdata <- subset(p105, Genotype=='En1cre')
WT_p105 <- newdata <- subset(p105, Genotype=='WT')
```

Subset p112 by Genotype
```{r}
SOD1_p112 <- newdata <- subset(p112, Genotype=='SOD1')
SOD1En1cre_p112 <- newdata <- subset(p112, Genotype=='SOD1En1cre')
En1cre_p112 <- newdata <- subset(p112, Genotype=='En1cre')
WT_p112 <- newdata <- subset(p112, Genotype=='WT')
```

Mean for each mouse

P49
```{r}
p49_means_mouse_by_Genotype <- p49 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

P56
```{r}
p56_means_mouse_by_Genotype <- p56 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

P63
```{r}
p63_means_mouse_by_Genotype <- p63 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

P70
```{r}
p70_means_mouse_by_Genotype <- p70 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p77
```{r}
p77_means_mouse_by_Genotype <- p77 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p84
```{r}
p84_means_mouse_by_Genotype <- p84 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p91
```{r}
p91_means_mouse_by_Genotype <- p91 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p98
```{r}
p98_means_mouse_by_Genotype <- p98 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p105
```{r}
p105_means_mouse_by_Genotype <- p105 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p112
```{r}
p112_means_mouse_by_Genotype <- p112 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

Dataframe with mean per mouse all timepoints

```{r}
df_mean_per_mouse <- rbind(p49_means_mouse_by_Genotype, p56_means_mouse_by_Genotype, p63_means_mouse_by_Genotype, p70_means_mouse_by_Genotype, p77_means_mouse_by_Genotype, p84_means_mouse_by_Genotype, p91_means_mouse_by_Genotype, p98_means_mouse_by_Genotype, p105_means_mouse_by_Genotype, p112_means_mouse_by_Genotype)
```


#Plots

Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the WT mice
```{r}
color_palette = c("#5B1A18","#FD6467",  "#F1BB7B")

```

#Genotype-wise analyses

##Plots used in paper (onset and p112)

Preparing the ggsignif function to use Dunnet's p-values
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}

```


one-way anova at p112

```{r}

#Ordering Genotype variables by manipulation so SOD1 becomes 'control' in Dunnet post-hoc tests
p112$Genotype <- factor(p112$Genotype, levels = c("SOD1", "WT", "En1cre","SOD1En1cre"))

#speed
anova_speed_1way_onset <- aov(Avg.Speed ~ Genotype, data = p112)
summary(anova_speed_1way_onset)

DunnettTest(x=p112$Avg.Speed, g=p112$Genotype)

#cadence
anova_cadence_1way_onset <- aov(MeanCadence ~ Genotype, data = p112)
summary(anova_cadence_1way_onset)

DunnettTest(x=p112$MeanCadence, g=p112$Genotype)


#stride
anova_stride_1way_onset <- aov(MeanStride ~ Genotype, data = p112)
summary(anova_stride_1way_onset)

DunnettTest(x=p112$MeanStride, g=p112$Genotype)


#acceleration
anova_acceleration_1way_onset <- aov(PeakAcc. ~ Genotype, data = p112)
summary(anova_acceleration_1way_onset)

DunnettTest(x=p112$PeakAcc., g=p112$Genotype)

```
Plots

```{r}

#speed
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
p112$Genotype <- factor(p112$Genotype, levels = c("SOD1", "SOD1En1cre", "WT"))
anova_speed_1way_onset <- aov(Avg.Speed ~ Genotype, data = p112)
summary(anova_speed_1way_onset)

Dunnet<- glht(anova_speed_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("WT", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= p112, aes(x=Genotype, y=Avg.Speed)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Arial"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
      geom_jitter(data=p112_means_mouse_by_Genotype, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112, position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 32) + ylab("Average speed (cm/s)") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
    ggsignif::geom_signif(family = "Arial",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(28, 22,25)
  )

#Acceleration
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
p112$Genotype <- factor(p112$Genotype, levels = c("SOD1", "SOD1En1cre", "WT"))
anova_acceleration_1way_onset <- aov(PeakAcc. ~ Genotype, data = p112)

Dunnet<- glht(anova_acceleration_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("WT", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= p112, aes(x=Genotype, y=PeakAcc.)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Arial"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
          geom_jitter(data=p112_means_mouse_by_Genotype, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112, position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 3.3)+ ylab("Peak acceleration (cm/s^2)") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
          ggsignif::geom_signif(family = "Arial",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(3.15,2.65, 2.9)
  )

#Cadence
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
p112$Genotype <- factor(p112$Genotype, levels = c("SOD1", "SOD1En1cre", "WT"))
anova_cadence_1way_onset <- aov(MeanCadence ~ Genotype, data = p112)

Dunnet<- glht(anova_cadence_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("WT", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= p112, aes(x=Genotype, y=MeanCadence)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+
  theme(text = element_text(family = "Arial"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
          geom_jitter(data=p112_means_mouse_by_Genotype, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112,position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 6)+ ylab("Step frequency (Hz)") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
        ggsignif::geom_signif(family = "Arial",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(5.5,4.7,5.1)
  )

#Stride
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
p112$Genotype <- factor(p112$Genotype, levels = c("SOD1", "SOD1En1cre", "WT"))
anova_stride_1way_onset <- aov(MeanStride ~ Genotype, data = p112)


Dunnet<- glht(anova_stride_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("WT", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= p112, aes(x=Genotype, y=MeanStride)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Arial"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
        geom_jitter(data=p112_means_mouse_by_Genotype, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112, position = position_jitter(width = .2),alpha=0.3)+
  ylim(0,10) + ylab("Stride length (cm)") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Arial",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(9,7.5,8.25)
  )


```

Analysis of onset


create subset of videos from onset of locomotor phenotype (when mouse could not keep up with treadmill belt speed of 20 cm/s). If mouse did not show onset at any point, onset is set to p112, except for WT and En1cre, they are set to p63. for mice where they have e.g had 2 trials at 20 cm and then one trial at 15 cm, the 'onset' has been counted as the next session where all trials were under 20 cm in speed.

triplicates
```{r}
class(df2$mouse_number)

onset_730 <-df2 %>% filter(mouse_number =='730' & time_point == 84)
onset_734 <- df2%>% subset((mouse_number =='734' & time_point == 84))
onset_736 <- df2%>% subset((mouse_number =='736' & time_point == 112))
onset_738 <- df2%>% subset((mouse_number =='738' & time_point == 91))
onset_788 <- df2%>% subset((mouse_number =='788' & time_point == 84))
onset_702 <- df2%>% subset((mouse_number =='702' & time_point == 98))
onset_615 <- df2%>% subset((mouse_number =='615' & time_point == 91))
onset_589 <- df2%>% subset((mouse_number =='589' & time_point == 112))
onset_607 <- df2%>% subset((mouse_number =='607' & time_point == 112))
onset_603 <- df2%>% subset((mouse_number =='603' & time_point == 112))

onset_wt <- df2%>% subset((Genotype == 'WT' & time_point == 63))
#onset_En1cre <- df2%>% subset((Genotype == 'En1cre' & time_point == 63))


#did not include En1cre as we dod not collect data from them until p112, so do not think they should be included for analysis
onset_all <- rbind(onset_730, onset_734, onset_736, onset_738, onset_788, onset_702, onset_615, onset_589, onset_607, onset_603, onset_wt)
```

Means


for mean
```{r}

onset_730 <-df_mean_per_mouse %>% filter(mouse_number =='730' & time_point == 84)
onset_734 <- df_mean_per_mouse%>% subset((mouse_number =='734' & time_point == 84))
onset_736 <- df_mean_per_mouse%>% subset((mouse_number =='736' & time_point == 112))
onset_738 <- df_mean_per_mouse%>% subset((mouse_number =='738' & time_point == 91))
onset_788 <- df_mean_per_mouse%>% subset((mouse_number =='788' & time_point == 84))
onset_702 <- df_mean_per_mouse%>% subset((mouse_number =='702' & time_point == 98))
onset_615 <- df_mean_per_mouse%>% subset((mouse_number =='615' & time_point == 91))
onset_589 <- df_mean_per_mouse%>% subset((mouse_number =='589' & time_point == 112))
onset_607 <- df_mean_per_mouse%>% subset((mouse_number =='607' & time_point == 112))
onset_603 <- df_mean_per_mouse%>% subset((mouse_number =='603' & time_point == 112))

onset_wt <- df_mean_per_mouse%>% subset((Genotype == 'WT' & time_point == 63))
#onset_En1cre <- df_mean_per_mouse%>% subset((Genotype == 'En1cre' & time_point == 63))


#did not include En1cre as we dod not collect data from them until p112, so do not think they should be included for analysis
onset_mean <- rbind(onset_730, onset_734, onset_736, onset_738, onset_788, onset_702, onset_615, onset_589, onset_607, onset_603, onset_wt)

```



onset only

```{r}

#speed
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "WT", "En1cre", "SOD1En1cre"))
anova_speed_1way_onset <- aov(Avg.Speed ~ Genotype, data = onset_all)
summary(anova_speed_1way_onset)

Dunnet<- glht(anova_speed_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("WT", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= onset_all, aes(x=Genotype, y=Avg.Speed)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
      geom_jitter(data=onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 32) + ylab("Average speed cm/s") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
    ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(23,26,29)
  )

#Acceleration
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "WT", "En1cre", "SOD1En1cre"))
anova_acceleration_1way_onset <- aov(PeakAcc. ~ Genotype, data = onset_all)

Dunnet<- glht(anova_acceleration_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("WT", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= onset_all, aes(x=Genotype, y=PeakAcc.)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
          geom_jitter(data=onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 3.3)+ ylab("Peak acceleration (cm/s^2)") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
          ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(2.65, 2.9,3.15)
  )

#Cadence
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "WT", "En1cre", "SOD1En1cre"))
anova_cadence_1way_onset <- aov(MeanCadence ~ Genotype, data = onset_all)

Dunnet<- glht(anova_cadence_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("WT", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= onset_all, aes(x=Genotype, y=MeanCadence)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
          geom_jitter(data=onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 6)+ ylab("Step frequency (Hz)") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
        ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(4.7,5.1,5.5)
  )

#Stride
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "WT", "En1cre", "SOD1En1cre"))
anova_stride_1way_onset <- aov(MeanStride ~ Genotype, data = onset_all)


Dunnet<- glht(anova_stride_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("WT", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= onset_all, aes(x=Genotype, y=MeanStride)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
        geom_jitter(data=onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0,9) + ylab("Stride length (cm)") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(6,7,8)
  )


```


one-way anova at onset

```{r}

#Ordering Genotype variables by manipulation so SOD1 becomes 'control' in Dunnet post-hoc tests
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "WT", "En1cre", "SOD1En1cre"))


#speed
anova_speed_1way_onset_all <- aov(Avg.Speed ~ Genotype, data = onset_all)
summary(anova_speed_1way_onset_all)
DunnettTest(x=onset_all$Avg.Speed, g=onset_all$Genotype)

#cadence
anova_cadence_1way_onset_all <- aov(MeanCadence ~ Genotype, data = onset_all)
summary(anova_cadence_1way_onset_all)
DunnettTest(x=onset_all$MeanCadence, g=onset_all$Genotype)

#stride
anova_stride_1way_onset_all <- aov(MeanStride ~ Genotype, data = onset_all)
summary(anova_stride_1way_onset_all)
DunnettTest(x=onset_all$MeanStride, g=onset_all$Genotype)

#acceleration
anova_acceleration_1way_onset_all <- aov(PeakAcc. ~ Genotype, data = onset_all)
summary(anova_acceleration_1way_onset_all)
DunnettTest(x=onset_all$PeakAcc., g=onset_all$Genotype) #Dunnet's test

```

LONGITUDINAL


##Statistics

Removing En1
```{r}
df2<-df2[!(df2$Genotype=='En1cre'),] 
```


For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the wt mice

Longitudinal plot Speed
```{r}
df2$time_point <-  as.numeric(df2$time_point)

ggplot(df2, aes(x=time_point, y= Avg.Speed, color=Genotype)) +
  geom_point(stat = "summary",fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary", fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Average speed (cm/s)")

```

Peak acceleration
```{r}
ggplot(df2, aes(x=time_point, y= PeakAcc., color=Genotype)) +
  geom_point(stat = "summary",fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary",fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Peak acceleration (cm/s^2)")

```

MeanCadence
```{r}
ggplot(df2, aes(x=time_point, y= MeanCadence, color=Genotype)) +
  geom_point(stat = "summary", fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary", fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Step frequency (Hz)")

```

MeanStride
```{r}
ggplot(df2, aes(x=time_point, y= MeanStride, color=Genotype)) +
  geom_point(stat = "summary",fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary",fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Stride length (cm)")

```


Two-way anova across timepoints
```{r}
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
df2$Genotype <- factor(df2$Genotype, levels = c("SOD1", "WT", "En1cre", "SOD1En1cre"))

df2$time_point <- as.factor(df2$time_point) #independent variables have to be recoded as factors to allow 'stratification' by timepoint for two-way anova

### Speed
anova_speed_2way <- aov(Avg.Speed ~ Genotype*time_point, data = df2)
summary(anova_speed_2way)

emmeans(anova_speed_2way,specs = trt.vs.ctrl~ Genotype|time_point)


### Cadence
anova_cadence_2way <- aov(MeanCadence ~ Genotype*time_point, data = df2)
summary(anova_cadence_2way)

emmeans(anova_cadence_2way,specs = trt.vs.ctrl~ Genotype|time_point)

### Stride
anova_stride_2way <- aov(MeanStride ~ Genotype*time_point, data = df2)
summary(anova_stride_2way)

emmeans(anova_stride_2way,specs = trt.vs.ctrl~ Genotype|time_point)


###Acceleration

anova_acceleration_2way <- aov(PeakAcc. ~ Genotype*time_point, data = df2)
summary(anova_acceleration_2way)

emmeans(anova_acceleration_2way,specs = trt.vs.ctrl~ Genotype|time_point)

```

