---
title: "WS analysis"
output:
  word_document: default
  html_document: default
date: '2023-03-14'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pacman)
p_load(tidyverse, ggsignif, multcomp, broom, emmeans, DescTools)
setwd("~/Documents/Systemic ESYT1 project locomotion, behavior, survival, anatomical data/Locomotion analysis cleaned/Weight support analysis/") 
df<- read.csv("WS_mean_new_method.csv", header=T)
colnames(df)[11]  <- "name"

#reverse so closer to back top is 1 and closer to front paw is 0

df$tb_ratio_mean_reverse <- df$tb_ratio_mean
df$tb_ratio_mean <- df$tb_ratio_mean_reverse-1
df$tb_ratio_mean<- abs(df$tb_ratio_mean)

```

```{r}
df2 <- df
df2$time_point <- 0
df2$time_point <- ifelse(grepl('p49_', df2$name), 49, df2$time_point)
df2$time_point <- ifelse(grepl('p56_', df2$name), 56, df2$time_point)
df2$time_point <- ifelse(grepl('p63_', df2$name), 63, df2$time_point)
df2$time_point <- ifelse(grepl('p70_', df2$name), 70, df2$time_point)
df2$time_point <- ifelse(grepl('p77_', df2$name), 77, df2$time_point)
df2$time_point <- ifelse(grepl('p84_', df2$name), 84, df2$time_point)
df2$time_point <- ifelse(grepl('p91_', df2$name), 91, df2$time_point)
df2$time_point <- ifelse(grepl('p98_', df2$name), 98, df2$time_point)
df2$time_point <- ifelse(grepl('p105_', df2$name), 105, df2$time_point)
df2$time_point <- ifelse(grepl('p112_', df2$name), 112, df2$time_point)

```

```{r}

df2$mouse_number <- 0

#SOD1En1cre
df2$mouse_number <- ifelse(grepl('736_', df$name), '736', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('589_', df$name), '589', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('603_', df$name), '603', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('607_', df$name), '607', df2$mouse_number)
#En1re
df2$mouse_number <- ifelse(grepl('561_', df$name), '561', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('694_', df$name), '694', df2$mouse_number) 
df2$mouse_number <- ifelse(grepl('696_', df$name), '696', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('698_', df$name), '698', df2$mouse_number)
#SOD1
df2$mouse_number <- ifelse(grepl('730_', df$name), '730', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('734_', df$name), '734', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('738_', df$name), '738', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('702_', df$name), '702', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('625_', df$name), '625', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('627_', df$name), '627', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('615_', df$name), '615', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('788_', df$name), '788', df2$mouse_number)
#WT
df2$mouse_number <- ifelse(grepl('692_', df$name), '692', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('700_', df$name), '700', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('706_', df$name), '706', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('708_', df$name), '708', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('553_', df$name), '553', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('555_', df$name), '555', df2$mouse_number)
```


Assign genotypes
```{r}
df2$Genotype <- 0


df2$Genotype <- ifelse(df2$mouse_number==702, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==730, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==734, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==738, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==625, 'SOD1', df2$Genotype) #consider deleting, SOD/WT injected with hEsyt1
df2$Genotype <- ifelse(df2$mouse_number==627, 'SOD1', df2$Genotype) #consider deleting, SOD/WT injected with hEsyt1
df2$Genotype <- ifelse(df2$mouse_number==615, 'SOD1', df2$Genotype) #consider deleting, SOD/En injected with GFP
df2$Genotype <- ifelse(df2$mouse_number==788, 'SOD1', df2$Genotype) #consider deleting, SOD/En injected with GFP
df2$Genotype <- ifelse(df2$mouse_number==700, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==692, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==706, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==708, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==553, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==555, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==694, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==698, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==696, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==561, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==736, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==589, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==603, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==607, 'SOD1En1cre', df2$Genotype)

```

Reduce dataset to the 6 SOD1 that were injected with GFP

```{r}
df2<-df2[!(df2$mouse_number==625 | df2$mouse_number==627),] 
```


```{r}
p49 <- newdata <- subset(df2, time_point==49)
p56 <- newdata <- subset(df2, time_point==56)
p63 <- newdata <- subset(df2, time_point==63)
p70 <- newdata <- subset(df2, time_point==70)
p77 <- newdata <- subset(df2, time_point==77)
p84 <- newdata <- subset(df2, time_point==84)
p91 <- newdata <- subset(df2, time_point==91)
p98 <- newdata <- subset(df2, time_point==98)
p105 <- newdata <- subset(df2, time_point==105)
p112 <- newdata <- subset(df2, time_point==112)
```


Compute mean for each mouse for each timepoint

```{r}
df_means_mouse_by_Genotype <- df2 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(tb_ratio_mean = mean(tb_ratio_mean)) %>% 
  arrange(Genotype)
```

Analysis of p112

```{r}
p112 <- subset(df_means_mouse_by_Genotype, time_point==112) #Mean values
p112_all <- subset(df2, time_point==112) #Video triplicates
```


p112 Analysis using triplicates
```{r}

p112_all$Genotype <- as.factor(p112_all$Genotype)
p112_all$Genotype <- factor(p112_all$Genotype, levels = c( "SOD1","SOD1En1cre", "WT", "En1cre"))

anova_tb_ratio <- aov(p112_all$tb_ratio_mean ~ Genotype, data = p112_all)
summary(anova_tb_ratio)

DunnettTest(x=p112_all$tb_ratio_mean, g=p112_all$Genotype)

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}

#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("WT", "En1cre", "SOD1", "SOD1En1cre"))
p112_all$Genotype <- factor(p112_all$Genotype, levels = c("WT", "En1cre", "SOD1", "SOD1En1cre"))

#color_palette inspired colorpalette
color_palette = c("#5B1A18","#FD6467",  "#F1BB7B")

Dunnet<- glht(anova_tb_ratio, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= p112_all, aes(x=Genotype, y=tb_ratio_mean)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  theme(text = element_text(family = "Arial"))+
  ylab("Weight support capability")+
    geom_jitter(data=p112, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(), text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Arial",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(0.8,0.70,0.75)
  )

```

Analysis of onset

one-way anova at onset

Analysis of onset


create subset of videos from onset of locomotor phenotype (when mouse could not keep up with treadmill belt speed of 20 cm/s). If mouse did not show onset at any point, onset is set to p112, except for WT and En1cre, they are set to p63. for mice where they have e.g had 2 trials at 20 cm and then one trial at 15 cm, the 'onset' has been counted as the next session where all trials were under 20 cm in speed.

triplicates
```{r}
class(df2$mouse_number)

onset_730 <-df2 %>% filter(mouse_number =='730' & time_point == 84)
onset_734 <- df2%>% subset((mouse_number =='734' & time_point == 84))
onset_736 <- df2%>% subset((mouse_number =='736' & time_point == 112))
onset_738 <- df2%>% subset((mouse_number =='738' & time_point == 91))
onset_788 <- df2%>% subset((mouse_number =='788' & time_point == 84))
onset_702 <- df2%>% subset((mouse_number =='702' & time_point == 98))
onset_615 <- df2%>% subset((mouse_number =='615' & time_point == 91))
onset_589 <- df2%>% subset((mouse_number =='589' & time_point == 112))
onset_607 <- df2%>% subset((mouse_number =='607' & time_point == 112))
onset_603 <- df2%>% subset((mouse_number =='603' & time_point == 112))

onset_WT <- df2%>% subset((Genotype == 'WT' & time_point == 63))
#onset_En1cre <- df2%>% subset((Genotype == 'En1cre' & time_point == 63))


#did not include En1cre as we dod not collect data from them until p112, so do not think they should be included for analysis
onset_all <- rbind(onset_730, onset_734, onset_736, onset_738, onset_788, onset_702, onset_615, onset_589, onset_607, onset_603, onset_WT)
```
Means

Means


for mean
```{r}

onset_730 <-df_means_mouse_by_Genotype%>% filter(mouse_number =='730' & time_point == 84)
onset_734 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='734' & time_point == 84))
onset_736 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='736' & time_point == 112))
onset_738 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='738' & time_point == 91))
onset_788 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='788' & time_point == 84))
onset_702 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='702' & time_point == 98))
onset_615 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='615' & time_point == 91))
onset_589 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='589' & time_point == 112))
onset_607 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='607' & time_point == 112))
onset_603 <- df_means_mouse_by_Genotype%>% subset((mouse_number =='603' & time_point == 112))

onset_WT <- df_means_mouse_by_Genotype%>% subset((Genotype == 'WT' & time_point == 63))
#onset_En1cre <- df_means_mouse_by_Genotype%>% subset((Genotype == 'En1cre' & time_point == 63))


#did not include En1cre as we dod not collect data from them until p112, so do not think they should be included for analysis
onset_mean <- rbind(onset_730, onset_734, onset_736, onset_738, onset_788, onset_702, onset_615, onset_589, onset_607, onset_603, onset_WT)

```

one-way anova at onset


```{r}

#Ordering Genotype variables by manipulation so SOD1 becomes 'control' in Dunnet post-hoc tests
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "WT", "En1cre", "SOD1En1cre"))

anova_tb_ratio <- aov(tb_ratio_mean ~ Genotype, data = onset_all)
summary(anova_tb_ratio)

DunnettTest(x=onset_all$tb_ratio_mean, g=onset_all$Genotype)

#Plots

#Re-ordering so the experimental groups appear in order of control level
onset_mean$Genotype <- factor(onset_mean$Genotype, levels = c("WT", "En1cre", "SOD1", "SOD1En1cre"))
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("WT", "En1cre", "SOD1", "SOD1En1cre"))

#color_palette inspired colorpalette
color_palette = c("#5B1A18","#FD6467",  "#F1BB7B")

Dunnet<- glht(anova_tb_ratio, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))


ggplot(data= onset_all, aes(x=Genotype, y=tb_ratio_mean)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  theme(text = element_text(family = "Arial"))+
  ylab("Weight support capability")+
    geom_jitter(data=onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=onset_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(), text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Arial",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(0.70,0.75,0.8)
  )

```